Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-09-20 paper/paper.Rmd
rewrite case study summer section

## Summary

The non-spatial CV estimate of AUC (0.76) is around 0.13 better compared to the spatial CV estimate of 0.64.
The better performance results from over-optimism in the estimation with non-spatial CV.
Spatial CV, in contrast, provides a better measure of a model's ability to generalize from the training sample --- in this case study, from the specific hillslopes and historical landslides in the training sample.
It is also expected that spatial CV results better reflect the performance of a model on new, spatially separate data.
The magnitude of the difference between spatial and non-spatial CV estimates may depend on the dataset, the strength on spatial or spatiotemporal autocorrelation, and the learner itself.
Algorithms with a higher tendency to overfit to the training set will tend to have a larger spread in such scenarios.

<!-- AB: TODO: Check if the following paragraphs overlaps with the Discussion section; perhaps merge into Discussion. -->
Since spatial CV can also be used for hyperparameter tuning in the \pkg{mlr3} framework, the model can be optimized to show an improved performance in a specific predictive setting [@schratz2019].
This may, for example, involve a reduced maximum tree depth in the Ecuador case study, since this leads to stronger generalization and reduced overfitting.

>>>>

## Interpretation

The non-spatial CV estimate of AUC (0.76) is around 0.13 higher compared to the spatial CV estimate of 0.64.
The higher non-spatial CV value results from over-optimism in the estimation with non-spatial CV and does imply better performance.
Spatial CV, in contrast, provides a better/more accurate measure of a model's ability to generalize from the training sample --- in this case study, from the specific hillslopes and historical landslides in the training sample.
It is also expected that spatial CV results reflect a more honest performance of a model on new, spatially separate data.
The magnitude of the difference between spatial and non-spatial CV estimates may depend on the dataset, the strength on spatial or spatiotemporal autocorrelation, and the learner itself.
Algorithms with a higher tendency to overfit to the training set will tend to have a larger spread in such scenarios.

<!-- AB: TODO: Check if the following paragraphs overlaps with the Discussion section; perhaps merge into Discussion. -->
Since spatial CV can also be used for hyperparameter tuning in the \pkg{mlr3} framework, the model can be optimized to show an improved performance in a specific predictive setting [@schratz2019].
This may, for example, involve a reduced maximum tree depth in the Ecuador case study, since this parameter leads to stronger generalization and reduced overfitting.
